Casual Bandits

### 2016-01-18 causal-bandit/icml-2015/setup.tex
More on parallel bandits.

\todom{Mention pruning of actions?}
We note that classical $K$-armed stochastic bandit problem can be recovered in our framework by considering a simple causal model with one edge connecting a single variable $X$ that can take on $K$ values to a reward variable $Y = r(X) + \eta$ for some arbitrary but unknown, real-valued function $r$ and a noise term $\eta$ that may depend on $X$.
The set of allowed interventions in this case is $\mathcal{A} = \{ do(X = k) \colon k \in \{1, \ldots, K\}\}$.
Furthermore, any causal bandit problem can reduced to a classical stochastic bandit problem by treating each possible intervention as an arm and ignoring all sampled values for the observed variables except for the reward.
Intuitively though, one would expect to perform better by making use of the extra observations.

@@

When not intervened upon, we assume that each $X_i \sim \bernoulli(q_i)$ where $\vec{q} = (q_1, \ldots, q_N) \in [0,1]^N$ so that $q_i = \P{X_i = 1}$.
The value of the reward variable is distributed as $Y = r(\vec{X}) + \eta$ where $r : \{0,1\}^N \to \R$ is an arbitrary, fixed, and unknown function and $\eta$ is $1$-subgaussian noise with a distribution that may depend on $\vec{X}$.
In the farming example, this choice of $Y$ models the fixed, noisy but unknown dependence of crop yield on the temperature, soil, and moisture.

% \eq{
% X_{t,i} \sim \begin{cases}
% \dirac(J_t) & \text{if } I_t = i \\
% \bernoulli(q_i) & \text{otherwise}\,.
% \end{cases}
% }
% where $\boldsymbol{q} \in [0,1]^N$ is a (possibly unknown) vector of probabilities with $q_i = \P{X_i = 1}$. 
% Note that if the learner did not choose an intervention ($I_t =  \bot$), then
% $X_{t,i} \sim \bernoulli(q_i)$ for all variables $i \in \set{1,\ldots,N}$.
% Finally the learner observes the reward $Y_t = r(X_t) + \eta_t$ where 
% \eq{
% r : \set{0,1}^N \to \R
% }
% is arbitrary (and unknown) 
% and $\eta_t$ is a $1$-subgaussian noise
% term (with the distribution possibly dependent on $X_t$). 
% In the farming example $X_{t,1}$ might represent whether the plants had sufficient water in the $t$th season (controllable by installing irrigation) 
% while $X_{t,2}$ might be whether or not there was a spring frost (controllable with heat-lamps or covers).

@@

% It is worth mentioning that the problem may be treated as a multi-armed bandit with $2N$ arms, one corresponding to each intervention.
% As we shall shortly see, this approach is usually not practical because the resulting algorithms do not exploit the structure in
% the covariates.

% \begin{remark}
% In order to be consistent with the literature on causal inference we use the notation $do()$ to denote the action of doing nothing and $do(X_{t,I_t} = J_t)$
% the action of intervening on the $I_t$th variable and setting it to equal $J_t$. 
% In the bandit community it is implicit that 
% algorithms selecting actions are intervening in the system. So it is sufficient to index actions according to the variable and value. 
% However, in causal graphs, it is essential to differentiate observing (or conditioning) on a variable taking a certain value, from 
% intervening to set that variable. Although in the specific causal graph we consider, observation and intervention are the same, we 
% deliberately introduce the do-notation \cite{Pearl2000} that makes this distinction clear so as to help provide a bridge between the 
% bandit and causal inference communities.
% \end{remark}

@@

>>>>

\todom{Mention pruning of actions?}
We note that classical $K$-armed stochastic bandit problem can be recovered in our framework by considering a simple causal model with one edge connecting a single variable $X$ that can take on $K$ values to a reward variable $Y = r(X) + \eta$ for some arbitrary but unknown, real-valued function $r$ and a noise term $\eta$ that may depend on $X$.
The set of allowed interventions in this case is $\mathcal{A} = \{ do(X = k) \colon k \in \{1, \ldots, K\}\}$.
Conversely, any causal bandit problem can reduced to a classical stochastic $|\mathcal{A}|$-armed bandit problem by treating each possible intervention as an arm and ignoring all sampled values for the observed variables except for the reward.
Intuitively though, one would expect to perform better by making use of the extra observations.

@@

In each round of the game, the learner selects a single variable to intervene on and the value to set it to while the remainder of the variables are simultaneously set by independently biased coin flips. 
The value of all variables are then used to determine the distribution of rewards for that round.
Formally, when not intervened upon we assume that each $X_i \sim \bernoulli(q_i)$ where $\vec{q} = (q_1, \ldots, q_N) \in [0,1]^N$ so that $q_i = \P{X_i = 1}$.
The value of the reward variable is distributed as $Y = r(\vec{X}) + \eta$ where $r : \{0,1\}^N \to \R$ is an arbitrary, fixed, and unknown function and $\eta$ is $1$-subgaussian noise with a distribution that may depend on $\vec{X}$.
In the farming example, this choice of $Y$ models the fixed, noisy but unknown dependence of crop yield on the temperature, soil, and moisture.

@@

@@

\begin{definition}
Let $\boldsymbol{q} \in [0,1]^N$. Then define $q^{(i)}$ to be the $i$th largest value of the set $\set{\min \set{q_j, 1 - q_j} : 1 \leq j \leq N}$.
For example
\eq{
\boldsymbol{q} &= (0,\, 2/3,\, 1/2,\, 1) \\
q^{(1)} &= 0,\,\, q^{(2)} = 0,\,\, q^{(3)} = 1/3,\,\, q^{(4)} = 1/2\,.
}
Define 
\eqn{
\label{def:m}
m(\boldsymbol{q}) = \min\set{1 \leq i \leq N: q^{(i+1)} \geq \frac{1}{i}}\,.
}
\end{definition}
